---
title: "Simulation Exercise: Exponential Distributions"
author: "Kenneth Fajardo"
date: "4/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This project investigates the variations of the exponential distribution as the population size approaches a large number $1000$. The probability distribution is given by
$$
f(x; \lambda) = 
\begin{cases} 
    \lambda e^{-\lambda x} & \text{if } x \geq 0 \\
    0         & \text{if } x < 0
\end{cases}
$$
where $\lambda$ is set to $0.2$. 

## Initialization
The data consists of the means of randomly generated probabilities using the *regexp(n, $\lambda$)* function, for a population size of 1000. We used a *for* loop for the calculation. $\lambda$ is set to $0.2$. For consistency, we also set the random seed to 1.
```{r init}
set.seed(1)
data <- c()     # create blank vector
n <- 1000       # population size
lambda <- 0.2   # assign lambda

for(i in seq_along(1:1000)){
    # calculate and store the mean of the values randomly generated 
    # over the exponential distribution 
    data <- rbind(data, mean(rexp(n, lambda)))
}
```

## 1. Show the sample mean and compare it to the theoretical mean of the distribution.
The theoretical mean $\mu$ in an exponential distribution is given by
$$
\mu = \frac{1}{\lambda}
$$
Substituting the value of $\lambda$, we get $\mu = 5$
```{r theoretical_mean} 
1/lambda
```
Now, for calculating the sample mean, we only need the *mean* function in R.
```{r sample_mean}
mean(data)
```

We see that **the sample mean is approximately equal to the theoretical mean** $(\mu \approx 5)$. To further illustrate, see the figure below.
```{r graph1}
hist(data, breaks=20, main="Illustration of the Mean in an Exponential Distribution", 
          xlab="Means of Exponential Distribution Values", ylab="Frequency")
abline(v=mean(data), col="red")
abline(v=1/lambda, col="green")
legend("topright", legend=c("Experimental","Theoretical"), col=c("red", "green"), lty=1)
```
## 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
The theoretical variance $\sigma^2$ in an exponential distribution is given by
$$
\sigma^2 = \frac{1}{\lambda^2 n}
$$

Substituting the value of $\lambda$, we get $\sigma^2 = 25$
```{r theoretical_var} 
1/(lambda^2 *n)
```

R has a built-in *var* function for calculating variance. Using it on our data, we get
```{r sample_var}
var(data)
```

We can see that the **theoretical variance is approximately equal to the sample variance** $(\sigma^2 \approx 0.025)$.

## 3. Show that the distribution is approximately normal.

One way to show normality is through histograms. Using the function *hist*, we can see the distribution of the means of exponential values. We, then, overlay the histogram with a bell curve, representing normal distribution.
```{r graph2}
h <- hist(data, breaks=20, main="Simulation of Exponential Distribution with n=1000", 
          xlab="Means of Exponential Distribution Values", ylab="Frequency")
xfit <- seq(min(data), max(data), length = 40) 
yfit <- dnorm(xfit, mean = mean(data), sd = sd(data)) 
yfit <- yfit * diff(h$mids[1:2]) * length(data) 

lines(xfit, yfit, col = "black", lwd = 2)
```

We can observe that the **frequencies of the means follow a normal distribution**. This is true according to the Central Limit Theorem when $n$ approaches large numbers.